### 2018-06-18 paper/paper.tex
Add initial, content-less version of paper

\documentclass[sigconf]{acmart}

\setcopyright{none}

\begin{document}

\title{Handwriting Recognition with SQL\\\hspace{6.88cm}\footnotesize and a tiny bit of web stuff}
\author{Noah Doersing}
\affiliation{}
\email{noah.doersing@student.uni-tuebingen.de}

\maketitle

\section*{Abstract}
This paper provides a sample of a \LaTeX\ document which conforms, somewhat loosely, to the formatting guidelines for ACM SIG Proceedings.\footnote{This is an abstract footnote}

\section{Introduction}

The main content goes here. testign

\section{Example/Demo?}

TODO really here? or something else?

\section{Implementation}

\subsection{Smoothing and Thinning}

test

\subsection{Curvature computation}

test

\subsection{Corner detection}

test

\subsection{Additional feature extraction}

test

\subsection{Mapping TODO}

test

\section{Evaluation}

test

\section{Related/Future work}

test

TODO conclusion, bibliography: essay, memo, memo on rand tablet, video of alan kay

\end{document}